### 2020-04-27 writeup/proof_sketch.tex
edits

\section{Proof Overview of Theorem \ref{thm_model_shift}}

\noindent\todo{A proof outline; including the following key lemma.}
To prove Theorem \ref{thm_cov_shift}, we study the spectrum of the random matrix model:
$$Q= \Sigma_1^{1/2}  Z_1^{\top} Z_1 \Sigma_1^{1/2}  + \Sigma_2^{1/2}  Z_2^{\top} Z_2 \Sigma_2^{1/2} ,$$
where $\Sigma_{1,2}$ are $p\times p$ deterministic covariance matrices, and $X_1=(x_{ij})_{1\le i \le n_1, 1\le j \le p}$ and $X_2=(x_{ij})_{n_1+1\le i \le n_1+n_2, 1\le j \le p}$ are $n_1\times p$ and $n_2 \times p$ random matrices, respectively, where the entries $x_{ij}$, $1 \leq i \leq n_1+n_2\equiv n$, $1 \leq j \leq p$, are real independent random variables satisfying
\begin{equation}\label{eq_12moment} %\label{assm1}
\mathbb{E} z_{ij} =0, \ \quad \ \mathbb{E} \vert z_{ij} \vert^2  = 1.
\end{equation}
%For now, we assume that the random variables $x_{ij}$ are i.i.d. Gaussian, but we know that universality holds for generally distributed entries. %have arbitrarily high moments,
%in the sense that for any fixed $k\in \mathbb N$, there is a constant $\mu_k>0$ such that
%\begin{equation}\label{eq_highmoment} %\label{eqn:subgaus}
%\max_{i,j}\left(\mathbb E|x_{ij}|^k\right)^{1/k} \le \mu_k n^{-1/2},  %\var \left(h_{xy}\right)^{1/2}
%\end{equation}
%for all $n$. %For simplicity, we assume that $k$ is a finite fixed integer, the strengths $d_1 > d_2 > \cdots > d_k >0$ are fixed constants, and $ \bu_i$,  $\bv_i$ are deterministic unit vectors.

\begin{lemma}\label{lem_cov_shift}
	In the setting of Theorem \ref{thm_model_shift}, we have with high probability $1-{\rm o}(1)$,
\begin{align}\label{lem_cov_shift_eq}
\tr ((w^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2) = \frac{1}{n_1+n_2}\cdot \tr \left( \frac{1}{a_1 M^\top M + a_2} \right) +\bigo{ n^{-1/2+\epsilon}},
\end{align}
	for any constant $\epsilon>0$. %, where $a_{3,4}$ are found using equations in  \eqref{m35reduced}.
	Moreover, when $n_1 = 0$, we have %we have that $a_1 = 0$ and $a_2 = (n_2-p) / n_2$, hence
	\begin{align}\label{lem_cov_shift_eq2} \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2} = \frac{p}{n_2-p} + \bigo{n^{-1/2+\epsilon}}, \end{align}
which is a well-known result for inverse Whishart matrices {\color{red}add some references}.
\end{lemma}
We will give the proof of this lemma in Section \ref{sec_maintools}.

\begin{lemma}\label{lem_minv}[\todo{ref?}]
	In the setting when $n_2 = c_2 p$ we have that as $p$ goes to infinity,
	\[ \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2} = \frac{p}{n_2 - p}. \]
	%{\color{blue} \[ \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2} = \frac{p}{n_2 - p} . \]}
	%\bigtr{\Sigma_2^{-1}}
\end{lemma}

\paragraph{A tighter bound for a special case.} We can get a bound tighter than Theorem \ref{thm_model_shift} as follows.

	\begin{proposition}\label{prop_model_shift_tight}
		In the setting of Theorem \ref{thm_model_shift}, assume that $\Sigma_1 =\id$,
		$\beta_t$ is i.i.d. with mean $0$ and variance $\kappa^2$ and $\beta_s - \beta_t$ is i.i.d. with mean $0$ and variance $d^2$.
		We set $\Delta_{\beta} = \bigbrace{(1 - \hat{w})^2 \kappa^2 + d^2)} \bigtr{Z}$
		and we have
		\begin{align*}
			\te(\hat{\beta}_t^{\MTL}) \le \te(\hat{\beta}_t^{\STL}) \text{ when: } & \Delta_{\vari} \ge \bigbrace{1 + \sqrt{\frac{p}{n_1}}}^4 \Delta_{\beta}, \\
			\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL}) \text{ when: } & \Delta_{\vari} \le \bigbrace{1 - \sqrt{\frac{p}{n_1}}}^4 \Delta_{\beta}.
		\end{align*}
	\end{proposition}