### 2020-06-04 writeup/experiments.tex
sec 3

\section{Experiments}

We connect our theory to pratical problems of interest.
First, we validate the single-task based metric on text and image classification tasks.
We show that for both datasets, single-task learning results can help predict positive or negative transfer in multi-task learning.
Second, we show that our proposed incremental training schedule improves the training efficiency of multi-task learning for predicting a particular target task.
%Third, we measure the data efficiency ratio of multi-task learning on six sentiment analysis tasks.
Finally, we validate the three components of our theory and measure the data efficiency of multi-task learning on text classification tasks.

\begin{table}[!b]
\begin{minipage}[t]{.40\textwidth}
	\centering
	\begin{tabular}{c c c}
		\toprule
		\multirow{2}{*}{{\bf Models}} & \multicolumn{2}{c}{\begin{minipage}{1.1in}\begin{center}
		MR, SST, SUBJ, CR, MPQA, TREC\end{center}\end{minipage}} \\
		\cmidrule(lr){2-3}
		& {\bf MTL} & {\bf IncTrain} \\
		\midrule
		{\bf MLP}  & 100\% & \textbf{xx\%} \\
		{\bf LSTM} & \textbf{36\%} & xx\% \\
		\bottomrule
		\end{tabular}
	\vspace{0.1in}
	\captionof{table}{Measuring the data efficiency ratio of muli-task learning.}
	\label{tab:taskonomy}
	\vspace{-0.2in}
\end{minipage}%
\quad
\begin{minipage}[t]{.58\textwidth}
	\centering
  \begin{tabular}{c c c c c}
	\toprule
		\multirow{2}{*}{{\bf Threshold}}  & \multicolumn{2}{c}{{\bf Sentiment
		analysis}} & \multicolumn{2}{c}{{\bf ChestX-ray14}} \\
		& Precision &  Recall & Precision &  Recall \\
		\cmidrule(lr){1-1} \cmidrule(lr){2-3} \cmidrule(lr){4-5}
		0.0 & 0.596 & 1.000 & 0.593 & 1.000 \\
		0.1 & \textbf{0.756} & \textbf{0.388} & \textbf{0.738} & \textbf{0.462} \\
		0.2 & 0.919 & 0.065 & 0.875 & 0.044 \\
		% 0.3 & 1.000 & 0.004 &     - &     - \\
	\bottomrule
	\end{tabular}
	\vspace{0.1in}
	\captionof{table}{Single-task learning results can help predict postive or negative transfer in multi-task learning.}
	\label{tab:mtl_better_than_stl}
	\vspace{-0.2in}
\end{minipage}
\end{table}

\subsection{Experimental Setup}

{\bf Datasets and models.} We describe the datasets and models we use in the experiments.

{\it Sentiment Analysis:} This dataset includes six tasks: movie review sentiment (MR), sentence subjectivity (SUBJ), customer reviews polarity (CR), question type (TREC), opinion polarity (MPQA), and the Stanford sentiment treebank (SST) tasks.

{For each task, the goal is to categorize sentiment opinions expressed in the text.
We use an embedding layer with GloVe embeddings
followed by an LSTM or MLP layer proposed by~\cite{lei2018simple}.
}
%multi-layer perceptron (MLP), LSTM, CNN on all tasks
%We use this task to verify our theoretical results on model capacity and task covariance in real world.

{\it ChestX-ray14:} This dataset contains 112,120 frontal-view X-ray images and each image has up to 14 diseases.
This is a 14-task multi-label image classification problem.
%We treat each label as one task a binary classification problem and formulate it as a 14-task multi-task learning problem.
%This dataset is curated where the labels
%We use the CheXNet model from~\cite{chexnet}, which is a 121-layer convolutional neural network on all tasks.

For all models, we share the main module across all tasks and assign a separate regression or classification layer on top of the shared module for each tasks.

\textbf{Training procedures and baselines.} We use round-robin training schedule for the MTL model.
We compare our incremental training scheduler of Algorithm \ref{alg_inc_train} to the round-robin training scheduler.
We focus on the prediction accuracy of a particular target task.

\subsection{Experimental Results}

\textbf{Improving multi-task training efficiency.}
We compare Algorithm \ref{alg_inc_train} to performing multi-task learning for predicting a particular target dataset.
Over six randomly selected task pairs from the sentiment analysis tasks, we find that Algorithm \ref{alg_inc_train} requires only $45\%$ of the computational cost to achieve the same performance on the target task, compared to the multi-task learning baseline.

\textbf{Improving transfer learning training efficiency.}
We show that Algorithm \ref{alg_inc_train} also applies to transfer learning settings.
Compared to fine-tuning the source model on the target task, we show that our proposed method reduces the computional cost by \alert{$xx\%$}, without sacrificing accuracy.

\textbf{MTL improves data efficiency ratio.}
We measure the data efficiency ratio on the sentiment analysis tasks.
In Table \ref{tab:taskonomy}, we find that by performing multi-task learning, only $39\%$ ({\cor or $36\%$?}) of the labeled data is needed to achieve comparable performance to single-task learning over all six tasks on LSTM.
The data efficiency ratio of using MLP is $100\%$ because the average performance of MTL is worse than the average of STL.
We further show that applying incremental training helps reduce the data efficiency ratio to \alert{$xx\%$}.
%If TREC is not included, we see that only $25\%$ of the labeled data is needed.

\begin{figure}[!t]
	\centering
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.975\textwidth]{figures/task_sim_norm_lstm.pdf}
		\caption{Task similarity}
		\label{fig_ab_sim}
	\end{subfigure}%
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.975\textwidth]{figures/ratio_norm_3_pairs_lstm.pdf}
		\caption{Data size}
		\label{fig_ab_data}
	\end{subfigure}
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.975\textwidth]{figures/ratio_alignment_norm_diff_all.pdf}
		\caption{Covariate shift}
		\label{fig_ab_cov}
	\end{subfigure}
	\caption{Validating the three takeaways of Section \ref{sec_insight} on sentiment analysis tasks. (a) Adding a semantically similar source task in MTL performs better than adding a dissimilar task.
	(b) As source/target data ratio increases, we observe a transition from positive to negative transfer.
	(c) As source/target data ratio increases, the performance gain from the covariance alignment procedure \cite{WZR20} over MTL increases for LSTM and MLP.}
	\label{fig_ablation}
	\vspace{-0.15in}
\end{figure}

\subsection{Ablation Studies}

\textbf{Explaining .} In Figure \ref{fig_ablation}, we validate the three components of our theory on the sentiment analysis dataset.
The experimental procedure is left to Appendix \ref{app_it}.

\textbf{Understanding MTL results via STL results.}
We show that STL results can be used to help understand MTL results.
We validate the single-task based metric proposed in Section \ref{sec_similarity} for predicting positive or negative transfer in MTL.
Table \ref{tab:mtl_better_than_stl} shows the result on both the sentiment analysis and the ChestX-ray14 datasets.
We find that using a threshold of $\tau = 0.1$, the STL results correctly predict positive or negative transfer with $75.6\%$ accuracy and $38.8\%$ recall among 150 task pairs.
We observe similar results for 91 task pairs from the ChestX-ray14 dataset.
The results show that STL performances are indicative of MTL performances.