### 2020-08-09 arxiv/main.tex

\documentclass{article}

%\usepackage[nonatbib]{neurips_2020}
\usepackage{fullpage}
\usepackage{subfloat}
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{macro_math}
\usepackage{macro_fan}
\usepackage{stmaryrd}
\usepackage{multirow}
\usepackage[us,12hr]{datetime}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage{caption,subcaption}
%\usepackage{breqn}

% Squished list environment to save space
\newcommand{\squishlist}{
\begin{list}{$\bullet$}
{ \setlength{\itemsep}{0pt}
\setlength{\parsep}{2pt}
\setlength{\topsep}{2pt}
\setlength{\partopsep}{0pt}
\setlength{\itemindent}{2pt}
\setlength{\leftmargin}{0.5cm}
% \setlength{\labelwidth}{1em}
}
}
\newcommand{\squishend}{\end{list}}

\setlength{\parskip}{0.3em}

\begin{document}

\title{Revisiting the Bias-Variance Tradeoff of Multi-Task Learning in High Dimensions}
\maketitle
%\date{{\ddmmyyyydate\today} at \currenttime}

\input{intro}
\input{setup}
\input{transfer}
%\input{extensions}
\input{experiments}
\input{related}

\bibliographystyle{unsrt}
\bibliography{rf,ref_mtl}

\appendix
\input{proofs_sec3}
\input{proof_main}
\input{proof_examples}
\input{proof_extension}
\input{proof_appd}
\input{app_exp}
\end{document}