### 2020-10-20 arxiv_full_version/main.tex

\documentclass{article}

%\usepackage[nonatbib]{neurips_2020}

\def\shownotes{0}
\usepackage{natbib}
\setcitestyle{authoryear,open={(},close={)}}
\renewcommand{\cite}{\citep}
\usepackage[export]{adjustbox}
\usepackage[margin=1in]{geometry}
%\usepackage{fullpage}
\usepackage{subfloat}
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{macro_math}
\usepackage{macro_fan}
\usepackage{stmaryrd}
\usepackage{multirow}
\usepackage[us,12hr]{datetime}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage{caption,subcaption}
%\usepackage{breqn}

\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}

%\setlength{\parskip}{0.3em}

%\title{Revisiting the Bias-Variance Tradeoff of Multi-Task Learning in High Dimensions}
\title{Sharp Bias-variance Tradeoffs of Hard Parameter Sharing in High-dimensional Linear Regression}
\usepackage{authblk}
\author[$\dagger$]{Hongyang R. Zhang\thanks{Equal contribution. Correspondence to hrzhang@northeastern.edu.}}
\author[$\ddagger$]{Fan Yang\samethanks}
\author[$\mathsection$]{Sen Wu}
\author[$\ddagger$]{Weijie S. Su}
\author[$\mathsection$]{Christopher R\'e}
\affil[$\dagger$]{Khoury College of Computer Sciences, Northeastern University}
\affil[$\ddagger$]{Department of Statistics, Wharton School, University of Pennsylvania}
\affil[$\mathsection$]{Department of Computer Science, Stanford University}

\begin{document}
\maketitle

\input{intro}
\input{setup}
%\input{examples}
\input{same_cov}
\input{diff_cov}
%\input{extensions}
%\input{algorithm}
\input{experiments}
\input{experiments_text}
\input{conclusions}
\bibliographystyle{plainnat}
\bibliography{rf,ref_mtl}
\appendix

\input{proof_same_cov}
\input{proof_gaussian_case}
\input{proof_appd}
\input{proof_examples}
\input{tools}
\end{document}