### 2021-06-12 revision_journal/backup/03:06:2021/manuscript.tex

\documentclass[aos,preprint]{imsart}
\usepackage{xr}
	\externaldocument{./suppl}

%% Packages
\def\shownotes{1}
%\usepackage{natbib}
%\setcitestyle{authoryear,open={(},close={)}}
%\renewcommand{\cite}{\citep}
\usepackage[export]{adjustbox}
%\usepackage[margin=1in]{geometry}
%\usepackage{fullpage}
\usepackage{subfloat}
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{macro_math}
\usepackage{macro_fan}
\usepackage{stmaryrd}
\usepackage{multirow}
\usepackage[us,12hr]{datetime}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage{caption,subcaption}
%\usepackage{breqn}

\usepackage{dsfont}

\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}

\endlocaldefs

\begin{document}

\begin{frontmatter}

\title{High-dimensional Asymptotics of Transferring from Different Domains in Linear Regression}

\runtitle{High-dimensional Asymptotics of Transferring from Different Domains}
%\thankstext{T1}{A sample additional note to the title.}

\begin{aug}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%Only one address is permitted per author. %%
%%Only division, organization and e-mail is %%
%%included in the address.                  %%
%%Additional information can be included in %%
%%the Acknowledgments section if necessary. %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author[A]{\fnms{Fan} \snm{Yang}\ead[label=e1]{}},
\author[B]{\fnms{Hongyang R.} \snm{Zhang}\ead[label=e2]{}},
\author[C]{\fnms{Sen} \snm{Wu}\ead[label=e3]{}},\\
\author[A]{\fnms{Weijie J.} \snm{Su}\ead[label=e4]{} },
\and
\author[C]{\fnms{Christopher} \snm{R\'e}\ead[label=e5]{}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Addresses                                %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\address{${ }^{1}$University of Pennsylvania, ${ }^{2}$Northeastern University, ${ }^{3}$Stanford University
}
%}\address[C]{
%}

\end{aug}

\input{intro}

\medskip
\noindent\textbf{Notations.}
%Let $\cE \define [\varepsilon_1, \varepsilon_2, \dots, \varepsilon_t] \in \real^{n \times t}$ denote the random noise.
%We can also write $Y = XB^{\star} + \cE$.
%Let $A = [A_1, A_2, \dots, A_t] \in \real^{r\times t}$ be a matrix notation that contains all the output layer parameters.
For an $n\times p$ matrix $X$, let $\lambda_{\min}(X)$ denote its smallest singular value and $\norm{X}$ denote its largest singular value.
Let $\lambda_1(X), \lambda_2(X), \cdots, \lambda_{p\wedge n}(X)$ denote the singular values of $X$ in decreasing order.
Let $X^+$ denote the Moore-Penrose psuedoinverse of $X$.
We refer to random matrices of the form $\frac {X^\top X} n$ as sample covariance matrices.
We say that an event $\Xi$ holds with high probability if the probability that $\Xi$ happens goes to $1$ as $p$ goes to infinity.
We use the big-O notation $g(n) = \OO(f(n))$ if there exists a constant $C$ such that $g(n) \le C \dot f(n)$ for large enough $n$. Moreover, we use the notation $g(n)\lesssim f(n)$ if $g(n) = \OO(f(n))$, and the notation $g(n)\sim f(n)$ if $g(n) \lesssim f(n))$ and $f(n) \lesssim g(n)$. In this paper, we will often write an identity matrix $\id_{n\times n}$ as $1$ without causing any confusions. 
%We shall use $\oo(1)$ to mean a small positive quantity that converges to 0 as $p$ goes to infinity.

\section{Main Results and Proof Techniques}\label{sec_HPS}
In this section, we define our model for the setting of transfer learning, and introduce several transfer learning estimators that will be considered in this paper.

\subsection{Assumptions}
Furthermore, we assume that the entries $z^{(1)}_{ij}$ and $z^{(2)}_{ij}$ have finite $\varphi$-th moment for some constant $\varphi>4$:
\be \label{conditionA2}
\mathbb{E} \vert z^{(1)}_{ij} \vert^\varphi  \le \tau^{-1},\quad \mathbb{E} \vert  z^{(2)}_{ij} \vert^\varphi   \le \tau^{-1}
\ee
for a small constant $\tau>0$. We assume that $\Sig^{(i)}$ has eigendecomposition
\be\label{eigen}
\Sig^{(i)}= O_i\Lambda_i O_i^\top, \quad  \Lambda_1=\text{diag}(\si_1^{(i)}, \ldots, \si^{(i)}_n),  \quad i=1,2,
\ee
where $O_i$ is the eigenmatrix and the eigenvalues satisfy that
\begin{equation}\label{assm3}
 \tau \le  \si^{(i)}_p \le\cdots\le \si^{(i)}_2 \le \si^{(i)}_1 \le \tau^{-1} \quad i=1,2.
\ee
%for a small constant $\tau>0$.

We assume that $\varepsilon^{(1)} \in \real^{n}$ and $\varepsilon^{(2)} \in \real^{n}$ are two independent random noise vectors with i.i.d entries of mean zero, variance $\sigma^2$, and bounded moment up to any order: for any fixed $k\in \N$, there exists a constant $C_k>0$ such that 
\be\label{eq_highmoments}
\E  | \ve^{(1)}_{i} |^k \le C_k,\quad \E  | \ve^{(2)}_{i} |^k \le C_k.
\ee
Finally, we assume that $\beta^{(1)}$ and $\beta^{(2)}$ are two arbitrary deterministic or random vectors that are independent of $X^{(1)}$, $X^{(2)}$, $\varepsilon^{(1)}$ and $\varepsilon^{(2)}$. 
%Moreover, we assume that 
%\be\label{eq_SNR}
%\tau\le \frac{\|\beta^{(1)}\|^2}{\|\beta^{(2)}\|^2} \le \tau^{-1},\quad  \frac{\|\beta^{(1)}\|^2}{\sigma^2} \ge \tau,
%\ee
%for a small constant $\tau >0$. The first condition means that the norms of the two model parameters $\beta^{(1)}$ and $\beta^{(2)}$ are comparable, while the second condition means that the singal-to-noise ratio cannot be too small.

In this paper, we consider the high-dimensional setting, where the sample ratios satisfy that  %We assume that the aspect ratios $d_1:= p/n_1$ and $d_2:=p/n_2$ satisfy that 
\be\label{assm2}
1+\tau \le \rho_1:=\frac{n_1}{p} \le p^{\tau^{-1}}, \quad 1+\tau \le \rho_2:=\frac{n_2}{p} \le p^{\tau^{-1}}, \quad \tau \le \frac{\rho_1}{\rho_2}\le \tau^{-1},
\ee
for a small constant $\tau >0$. If $\rho_i>p^{\tau^{-1}}$, $i=1,2$, we are basically in the low-dimensional region, where the law of large numbers and central limit theorem already  give good enough results without using the theory developed in this paper. The lower bounds $\rho_1> 1+\tau$ and ${\rho_2 > 1+ \tau}$ 
are to ensure that the sample covariance matrices $  (X^{(1)})^\top X^{(1)} $ and $  (X^{(2)})^\top X^{(2)} $ are non-singular with high probability, so that the ordinary least squares (OLS) estimator is well-defined for the linear regression problem on each task.

We summarize our basic assumptions here for future reference.
\begin{assumption}\label{assm_big1}
Let $\tau$ be a small constant. 
\begin{enumerate}
\item  $X^{(1)}$ and $X^{(2)}$ take the form \eqref{XofZ}, where $Z^{(1)}$ and $Z^{(2)}$ are respectively $n_1\times p$ and $n_2\times p$ random matrices with i.i.d. entries satisfying (\ref{assm1}) and \eqref{conditionA2}, $\Sig^{(1)}$ and $\Sig^{(2)}$ are deterministic positive definite symmetric matrices satisfying \eqref{eigen} and \eqref{assm3}. %and (\ref{assm2}). We assume that $T$ is an $M\times M$ deterministic diagonal matrix satisfying (\ref{simple_assumption}) and (\ref{assm3}).

\item $\varepsilon^{(1)} \in \real^{n}$ and $\varepsilon^{(2)} \in \real^{n}$ are random vectors independent from $X^{(1)}$ and $X^{(2)}$, and with i.i.d entries of mean zero, variance $\sigma^2$, and bounded moments as in \eqref{eq_highmoments}.

\item $\beta^{(1)}$ and $\beta^{(2)} $ are independent of $X^{(1)}$, $X^{(2)}$, $\varepsilon^{(1)}$ and $\varepsilon^{(2)} $. %, and satisfy \eqref{eq_SNR}.

\item $\rho_{1}$ and $\rho_{2}$ satisfy \eqref{assm2}.

\end{enumerate}
\end{assumption}

%\subsection{Transfer learning estimators}

%\FY{Introduce some popular transfer learning estimators: total loss estimator, and model averaging estimator.}

In this paper, we measure the performance of $\hat{\beta}_2^{\MTL} (\hat a)$ through the out-of-sample predication loss (test error). Consider a test data point $(x,y)$ generated from the same model as task 2: $y= x^\top \beta^{(2)} + \ve$, where $x \in \R^p$ and $\ve\in \R$ are independent of $X^{(1)}$, $X^{(2)}$, $\varepsilon^{(1)}$ and $\varepsilon^{(2)}$, and only $x$ is observable. We want to use $x^\top \hat{\beta}_2^{\MTL} (\hat a) $ to predict $y$, and we measure the predication loss using the mean squared error
$$\exarg{x}{\left\|y-x^\top \hat{\beta}_2^{\MTL} (\hat a) \right\|^2}= \left\|(\Sigma^{(2)})^{1/2} \left(\hat{\beta}_2^{\MTL} (\hat a)  - \beta^{(2)}\right)\right\|^2 + \sigma^2.  $$ 
Since $\sigma^2$ is a constant that does not depend on the model, we ignore it and define the predication loss as
\be\label{HPS_loss}
L(\hat{\beta}_2^{\MTL}(\hat a)):= \left\|(\Sigma^{(2)})^{1/2} \left(\hat{\beta}_2^{\MTL} (\hat a)  - \beta^{(2)}\right)\right\|^2 .
 \ee
We will compare it to the out-of-sample predication losses of the baseline estimators, $L(\hat{\beta}_2^{\STL})$ and $L(\hat{\beta}^{\AV}(\lambda))$ for some properly chosen $\lambda\in [0,1]$.

%$L(\hat{\beta}_2^{\STL})$ of the single-task OLS estimator  as a baseline.

Plugging \eqref{HPS_est} into \eqref{HPS_loss}, we get
\begin{align}
L(\hat{\beta}_2^{\MTL}(\hat a))  =  &\left\| (\Sigma^{(2)})^{1/2}\hat \Sigma(\hat a)^{-1} (X^{(1)})^\top X^{(1)} (\hat a\beta^{(1)}-\hat a^2\beta^{(2)}) \right. \nonumber\\
&\left. +(\Sigma^{(2)})^{1/2}\hat \Sigma(\hat a)^{-1} \left[ (X^{(2)})^\top \epsilon^{(2)} + \hat a  (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2.\label{Lbeta_HPS}
\end{align}
Note that if we replace $\hat a$ with a fixed number $a$ in \eqref{Lbeta_HPS}, then taking the expectation over $\ve^{(1)}$ and $\ve^{(2)}$, we can get a clean bias-variance decomposition of the predication loss:
\be\label{partial-L} \exarg{\ve^{(1)},\ve^{(2)}}{L(\hat{\beta}_2^{\MTL}(a)) } = L_{\bias} (a) + L_{\var}(a),\ee
where 
\be\label{Lbias} L_{\bias} (a) := \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma(a)^{-1} (X^{(1)})^\top X^{(1)} \left(a\beta^{(1)}- a^2\beta^{(2)}\right) \right\|^2 \ee
is called the bias term, which depends on the model bias between task 1 and task 2, and
\be\label{Lvar}  L_{\var}(a):=\sigma^2  \bigtr{{\Sigma^{(2)}\hat \Sigma(a)^{-1}  }} \ee
is called the variance term, which depends on the noise variance. Using concentration of the noise vectors $\ve^{(1)}$ and $\ve^{(2)}$, we can show that $L(\hat{\beta}_2^{\MTL}(a))$ is close to \eqref{partial-L} up to a small error as in the next lemma. In this paper, we say that an event $\Xi$ holds \emph{with high probability} (w.h.p.) if $\mathbb P(\Xi)\to 1$ as $p\to \infty$.
%predication loss for the HPS and OLS estimators.

 \begin{lemma}\label{lem_HPS_loss}
 Under Assumption \ref{assm_big1}, for any small constant $\e>0$ and large constant $C>0$, there exists a high probability event $\Xi$, on which the following estimates hold uniformly in $a\in \R$ and $\lambda\in [0,1]$: 
\begin{align}
L(\hat{\beta}_2^{\MTL}(a)) &=\left[ 1+ \OO(p^{-1/2+c})\right]\cdot \left[L_{\bias} (a) + L_{\var}(a)\right] \nonumber\\
&+ \OO \left[ p^{-C}\left( \|\beta^{(1)} \|^2  +  \|\beta^{(2)} \|^2\right)   \right], \label{L_HPS_simple}
\end{align}
\begin{align}
L(\hat{\beta}^{\AV}(\lambda) ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \lambda^2 \left\| (\Sigma^{(2)})^{1/2}(\beta^{(1)}-\beta^{(2)})\right\|^2 \nonumber \\
&+  \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2   \lambda^2 \bigtr{\Sigma^{(2)} [(X^{(1)})^\top X^{(1)}]^{-1}  }\nonumber\\
& + \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2  (1-\lambda)^2 \bigtr{\Sigma^{(2)} [(X^{(2)})^\top X^{(2)}]^{-1}  }  . \label{L_AVE_simple}
\end{align}
and 
\begin{align}
L(\hat{\beta}_2^{\STL} ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2  \bigtr{\Sigma^{(2)} [(X^{(2)})^\top X^{(2)}]^{-1}  }  . \label{L_STL_simple}
\end{align}
%Moreover, there exists a constant $C_0>0$ such that with high probability,
%\be\label{loss_large}
%g(a) > g(0) \quad \text{for all $a$ such that }|a|\ge C_0.
%\ee
%with high probability for any small constant $\e>0$. 
\end{lemma}

Since \eqref{L_HPS_simple} holds uniformly for all $a\in \R$, we can also apply it to $\hat{\beta}_2^{\MTL}(\hat a)$, where $\hat a$ is a random variable that may depend on $\ve^{(1)}$ and $\ve^{(2)}$. In practice, the parameter $ a$  is up to one's choice and may not be the global minimizer, so we have stated the result for an completely arbitrary $a$. (In fact, the optimization objective $g(a)$ is generally non-convex, so in some cases one can only find a local minimizer.) For a fixed $a\in \R$, the proof of \eqref{L_HPS_simple} is based on the sharp concentration bounds in Lemma \ref{largedeviation} of the supplement \cite{MTL_suppl}. To extend uniformly to all $a\in \R$, we will use a standard $\ve$-net argument, which leads to a small error $ p^{-C}\left( \|\beta^{(1)} \|^2  +  \|\beta^{(2)} \|^2\right) $. Note that this error is negligible unless $\beta^{(1)}$ and $ a\beta^{(2)}$ cancel each other almost exactly, and the noise variance $\sigma$ is very small. The proof of Lemma \ref{lem_HPS_loss} will be given in Appendix  \ref{app_firstpf} of the supplement \cite{MTL_suppl}.

%In this section, we show that the prediction loss of hard parameter sharing admits a clean bias-variance decomposition, when all tasks have the same features.

For the single-task predication loss \eqref{L_STL_simple}, we can calculate its asymptotic limit exactly using the following classical result in multivariate statistics.

\begin{lemma}[Theorem 2.4 of \cite{isotropic} and Theorem 3.14 of \cite{DY}]\label{fact_tr}
	%Let $X  \in \real^{n\times p}$ be a random matrix that satisfies Assumption \ref{assume_rm}.
	%Let $\Sigma\in\real^{p\times p}$ denote the population covariance matrix of $X$.
	Under Assumption \ref{assm_big1}, we have that for any deterministic $p\times p$ matrix $\Sigma$,
		\[ \bigtr{\Sigma \frac1{(X^{(2)})^\top X^{(2)} }  } %= \bigtr{ \frac1{(Z^{(2)})^\top Z^{(2)} }  } 
		= \frac{1}{n_2 - p}\bigtr{\Sigma (\Sigma^{(2)})^{-1}} +  \OO\left( \frac{p^{\e}}{\sqrt{n_2 p}}\cdot \frac{p}{n_2}\|A\| \right) \]
		with high probability for any small constant $\e>0$.
\end{lemma}
 %The above result has a long history in random matrix theory.
If the entries of $Z^{(2)}$ are i.i.d. Gaussian, then this result follows from the classical result for %the mean of 
the inverse Wishart distribution \cite{anderson1958introduction}. For a general non-Gaussian $Z^{(2)}$, this result can be obtained using the well-known Stieltjes transform method (cf. Lemma 3.11 of \cite{bai2009spectral}). Here we have presented the results from \cite{isotropic,DY}, which give an almost sharp convergence rate. With Lemma \ref{fact_tr}, we obtain that w.h.p.,
\begin{align}
L(\hat{\beta}_2^{\STL} ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \frac{p \sigma^2 }{n_2-p}   , \label{L_STL_simple01}
\end{align}
and
\begin{align}
L(\hat{\beta}^{\AV}(\lambda) ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \lambda^2 \left\| (\Sigma^{(2)})^{1/2}(\beta^{(1)}-\beta^{(2)})\right\|^2 \nonumber \\
&+  \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2 \left[\frac{ \lambda^2}{n_1-p}\bigtr{\Sigma^{(2)}(\Sigma^{(1)})^{-1}} + \frac{(1-\lambda)^2p  }{n_2-p}\right] . \label{L_AVE_simple01}
 \end{align}

On the other hand, much less is known about the predication loss of the HPS estimator. In particular, its exact dependence on the model parameters, including the model bias, noise variance, sample sizes, and population covariance matrices, is not well-understood so far. %\FY{What is known so far about this topic?} 
In Section \ref{sec_HPS_loss}, we give a rigorous analysis of the bias term \eqref{Lbias} and variance term \eqref{Lvar} in two different settings: (i) independent $X^{(1)}$ and $X^{(2)}$ with different sample sizes and same population covariance matrices $\Sigma^{(1)}=\Sigma^{(2)}$; (ii) independent $X^{(1)}$ and $X^{(2)}$ with different sample sizes and different population covariance matrices. In case (i), we provide the exact asymptotics for both the bias and variance terms, together with almost sharp convergence rates. We will use these results to illustrate the effects of the \emph{bias-variance tradeoff} and \emph{varying sample sizes}. In case (ii), we will prove an exact asymptotic variance limit, and employ the result to show the intricate effect of the covariate shift. On the other hand, in case (ii), we will provide an estimate on the bias term, which becomes exact only when $n_1\gg p$. %Here case (ii) is actually more general than (i) 
%In particular, we will use our results to analyze the effect of the following factors: \emph{bias-variance tradeoff}, \emph{sample sizes}, and \emph{covariate shift}.   
In Section \ref{sec_same}, we will consider a simple extension to a multi-task learning setting with more than two tasks and  the same covariates for all tasks. We will give the asymptotic limit of the predication loss for the HPS estimator, which provides useful insights into the more rich multi-task learning theory. Due to length constraint, we will explore the multi-task learning setting in greater details in other works.

%{\cor discuss about the bias-variance trade-off}
%However, to the best of our knowledge, the asymptotic limit of $L(\hat{\beta}_2^{\MTL}(a))$ has never been calculated exactly in any high-dimensional setting.

%\section{Predication loss for HPS estimator}\label{sec_HPS_loss}

\subsection{Covariate shift}\label{sec_diff}

%The previous section assumes that all tasks have the same sample size and feature vectors.
In this section, we consider the most general setting, where the feature covariates $X^{(1)}$ and $X^{(2)}$ of the two tasks have both different sample sizes and different population covariance matrices.
%The different covariates case differs from the same covariates case in two aspects.
%First, different tasks may have different sample sizes. In extreme scenarios, one task may have much less labeled data compared to another task.
In particular, the fact that population covariance matrices differ across tasks is often called ``covariate shift''. %which is characterized by the matrix $(\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}$. 
In this section, we describe the exact asymptotic variance limit in the high-dimensional setting, while the bias limit is much more complicated, and we can only give an estimate on it.

\iffalse
Unlike the previous section, we can no longer characterize the global minimum of $f(A, B)$.
This is because $f(A, B)$ is in general non-convex.
Instead, our result implies sharp bias-variance tradeoffs for any \emph{local minimizer} of $f(A, B)$.
We focus on the two-task case to better understand the impact of having different sample sizes and different covariates.
Let $n_1, n_2$ denote task one  and two's sample size, respectively.
Suppose
\begin{align*}
	X^{(1)} = Z^{(1)}(\Sigma^{(1)})^{1/2} \in \real^{n_1 \times p} \text{ and }
	X^{(2)} = Z^{(2)}(\Sigma^{(2)})^{1/2} \in \real^{n_2 \times p},
\end{align*}
where the entries of $Z^{(1)}$ and $ Z^{(2)}$ are drawn independently from a one dimensional distribution with zero mean, unit variance, and constant $\varphi$-th moment for a fixed $\varphi > 4$. $\Sigma^{(1)}\in \R^{p\times p}$ and $\Sigma^{(2)}\in \R^{p\times p}$ denote the population covariance matrices of task 1 and task 2, respectively.

Without loss of generality, we consider task two's prediction loss and the same result applies to task one.
We consider the case of $r = 1 < t = 2$, since when $r > 1$, the global minimum of $f(A, B)$ reduces to single-task learning (cf. Proposition 1 of \cite{WZR20}).
When $r = 1$, $B$ is a vector and $A_1, A_2$ are both scalars.
To motivate our study, we consider a special case where $A_1=A_2=1$.
Hence the HPS estimator is equal to $B$.
%Hence we can write down a closed form equation for any local minimizer of $f(A, B)$.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the estimator for task two as follows:
\begin{align}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \text{ where }
	\hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}. \label{def hatsig}
\end{align}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariance matrices, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.

To derive the bias and variance equation, we consider the expected loss conditional on the covariates as follows (the empirical loss is close to this expectation as will be shown in equation \eqref{claim_largedev2}):
 %similar to Claim \ref{claim_pred_err}
\begin{align}
	 \exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}}
	=& \bignorm{{\Sigma^{(2)}}^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2 \label{eq_bias_2task} \\
	& + \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}. \label{eq_variance_2task}
\end{align}
Equations \eqref{eq_bias_2task} and \eqref{eq_variance_2task} correspond to the bias and variance of HPS for two tasks, respectively.
\fi

%Our key result characterizes the asymptotic limit of the inverse of the sum of two arbitrarily different sample covariance matrices.
Compared to the results in Section \ref{sec_sizeratio}, the spectrum of $\hat{\Sigma}(a)^{-1}$ now not only depends on the sample sizes of both tasks, but also depends on the ``misalignment'' between $\Sigma^{(1)}$ and $\Sigma^{(2)}$. 
%However, capturing this intuition quantitatively turns out to be technically challenging.
%The main technical challenge of our result deals with the ``covariate shift'' between tasks one and two.
To capture this misalignment quantitatively, we introduce the covariate shift matrix %(rescaled by $a$) 
$$ M(a) \define a(\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}.$$ 
%Let $U\Lambda V^\top$ denote the SVD of $M$ and
Let $\lambda_1(a)\ge \lambda_2(a)\ge \dots\ge \lambda_p(a)$ be the singular values of $M$ in descending order. By \eqref{assm3}, these singular values satisfy that
\begin{equation}\label{assm32}
|a| \tau \le  \lambda_p(a) \le \cdots \le \lambda_2(a)\le \lambda_1(a) \le |a|\tau^{-1}. %, \quad \max\left\{\pi_A^{(n)}([0,\tau]), \pi_B^{(n)}([0,\tau])\right\} \le 1 - \tau .
\end{equation}
%for some small constant $0<\tau<1$. 
The main result of this section is the following theorem on the variance limit, which characterizes  the exact dependence of $L_{\var}(a)$ on the singular values of $M$.

\begin{theorem}\label{thm_main_RMT}
%	Let $c_{\varphi}$ be any fixed value within $(0, \frac{\varphi - 4}{2\varphi})$.
%	Assume that: a) the sample sizes $n_1$ and $n_2$ both satisfy Assumption \ref{assume_rm};
%	b) $M$'s singular values are all greater than $\tau$ and less than $1/\tau$;
%	c) task one's sample size is greater than $\tau p$ and task two's sample size is greater than $(1 + \tau) p$.
%	With high probability over the randomness of $X^{(1)}$ and $X^{(2)}$, we have the following limits:
Under Assumption \ref{assm_big1}, for any small constant $c>0$, there exists a high probability event $\Xi$, on which the following estimate holds for $L_{\var}(a)$ in \eqref{Lvar}: 
%equation \eqref{eq_variance_2task} $\tr[\Sigma^{(2)} \hat{\Sigma}^{-1}]$ (leaving out $\sigma^2$) satisfies the following estimate:
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{L_{\var}(a)- \frac{\sigma^2}{n_1+n_2}\bigtr{  \frac{1}{a_1 M(a)^\top M(a) + a_2  }  }}
				\le \frac{(n_1+n_2)^{2/\varphi + c}}{p^{1/2} (n_1+n_2)^{1/2}}\cdot\frac{p \sigma^2}{n_1+ n_2}  ,
			\end{align}
			 uniformly in all $a\in \R$. Here $(a_1,a_2)$ is the solution of the following system of equations
			\begin{align}
				a_1 + a_2 = 1- \frac{p}{n_1 + n_2}, \quad
				a_1 + \frac1{n_1 + n_2}  \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}

\end{theorem}

 With \eqref{assm32}, it is easy to see that 
$$ \frac{\sigma^2}{n_1+n_2}\bigtr{  \frac{1}{a_1 M(a)^\top M(a) + a_2  }  } \sim \frac{p \sigma^2}{n_1+ n_2}.$$
Hence the right-hand side of \eqref{lem_cov_shift_eq} is much smaller than this main term by a factor of $p^{-1/2} (n_1+n_2)^{-1/2+2/\varphi + c}$, which we believe to be sharp up to the $(n_1+n_2)^c$ factor. Lemma \ref{fact_tr} can be also regarded as a special case of Theorem \ref{thm_main_RMT}. %also extends  to the inverse of the sum of two sample covariance matrices.
To see this, when $M(a)=0$, we solve equation \eqref{eq_a12extra} to obtain that 
$$a_1 = \frac{n_1}{n_1+n_2},\quad a_2 = \frac{n_2-p}{n_1+ n_2},$$ 
and plug them into \eqref{lem_cov_shift_eq}.  The proof of Theorem \ref{thm_main_RMT} is based on some recent developments in random matrix theory \cite{Anisotropic}, and will be given in Appendix \ref{appendix RMT} of the supplement \cite{MTL_suppl}.
%For general $A_1,A_2$ that are not equal to one, we can still apply our result by rescaling $X^{(1)}$ and $M$ with $A_1 / A_2$.
%We defer a proof sketch of Theorem \ref{thm_main_RMT} until the end of the section.
%This amounts to replacing $M$ with $\frac{A_1}{A_2}M$ in Theorem \ref{thm_main_RMT}.

\subsection{The case of isotropic covariances} \label{sec_sizeratio}

 In this subsection, we consider the setting where $X^{(1)}$ and $X^{(2)}$ are independent and have the same population covariance matrices $\Sigma^{(1)}=\Sigma^{(2)}$. However, the two tasks can have different sample sizes $n_1\ne n_2$. In this case, we can obtain the exact asymptotic limits of the bias term \eqref{Lbias} and the variance term \eqref{Lvar}. We will use these results to illustrate the effects of the bias-variance tradeoff and varying sample sizes on the performance of the $\MTL$ estimator.

\begin{theorem}\label{cor_MTL_loss}
Under Assumption \ref{assm_big1}, suppose that $\Sigma^{(1)}=\Sigma^{(2)}$ and the entries of $Z^{(1)}$ and $Z^{(2)}$ are i.i.d. Gaussian random variables. Then for any small constant $\e>0$ and large constant $C>0$, there exists a high probability event $\Xi$, on which the following estimates hold uniformly in all $a\in \R$:
\be\label{Lvar_samplesize}
L_{\var}(a)= \sigma^2  \left[ \cal L_1(a)+ \OO\left({p^{c}}/{n_1}\right)\right]  ,
\ee
and
\begin{align}
L_{\bias}(a)&= \left[ \cal L_2(a)+  \OO\left(p^{-1/2+c}\right)\right] \cdot  \|\beta^{(1)}-a\beta^{(2)}\|^2 \nonumber\\
&+ \OO \left[ p^{-C}\left( \|\beta^{(1)} \|^2  +  \|\beta^{(2)} \|^2\right)   \right].\label{Lbias_samplesize}
	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le
	%   \left|L(\hat{\beta}_2^{\MTL}) - \frac{2d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} -\frac{\sigma^2 p}{n_1 + n_2 - p}  \right|
	% \le \varepsilon \cdot \frac{2d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} +  \OO(p^{-c/2}).\label{cor_MTL_error}
	%\left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] d^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \\
	%& +C \left[(p^{-c_\varphi}+p^{-c_\infty/2})(\sigma^2 +d^2)+p^{-c_\infty}\kappa^2 + %\frac{d^4+\sigma^2 d^2}{\kappa^2}\right],\nonumber
	 \end{align}
Here we define the functions
\begin{align*}
&\cal L_1(a):= \frac{2p}{  {n_2} (1-\xi_2) + a^2 n_1 (1-\xi_1)+ \sqrt{[ {n_2} (1-\xi_2) +a^2 n_1 (1-\xi_1)]^2 + 4a^2{n_2}{n_1} (\xi_1+\xi_2 -\xi_1\xi_2)}}, \\
&\cal L_2(a): = \frac1{a^2}\cdot \frac{1- 2\frac{\cal L_1(a)}{\xi_2[1 + \cal L_1(a)]} + \kappa(a)}{1- \xi_2 \kappa(a)},
\end{align*}
where we abbreviate $\xi_1:=p/n_1$, $\xi_2:=p/n_2$ and
\begin{align*}
\kappa(a):= \frac{\cal L_1(a)^2}{\xi_2^2[1+\cal L_1(a)]^2}  \left[1 - \frac{a^4 \cal L_1(a)^2 }{\xi_1[1+a^2\cal L_1(a) ]^2}\right]^{-1}.
%f_2(a):= \frac{a^4 \cal L_1(a)^2}{\xi_1^2 } \left[1 - \frac{a^4 \cal L_1(a)^2 }{\xi_1[1+a^2\cal L_1(a) ]^2}\right]^{-1},\\
%f_3(a):= \frac1{a^2}\frac{\xi_1}{\xi_2 +\cal L_1(a)}
 \end{align*}
%	In the setting of Example \ref{ex_sample_ratio}, assume that
%	%a) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$, b) $\Sigma_1=\Sigma_2=\id_p$, and c) %there exists a small constant $c_0>0$ such that
%	(i) both tasks sample sizes are at least $3p$;
%	(ii) noise variance is smaller than the shared signal variance: $\sigma^2 \lesssim  \kappa^2$;
%%	\be\label{choiceofpara0}
%%	p^{-1/2+c_0}\sigma^2 + p^{c_0}d^2\le \kappa^2\le p^{1-c_0} (\sigma^2 +d^2)  .  	\ee
%	%\be\label{choiceofpara0}
%%	(ii) the task-specific variance of $\beta_i$ is much smaller than the signal strength {\color{red}$d^2 = \oo( {\kappa^2})$}; \HZ{what does $\ll$ mean exactly?}
%%	(iii) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$.
%	(iii) task-specific variance is much smaller than the shared signal variance: $d^2 \le p^{-\e}{\kappa^2}$ for a small constant $c>0$.
%	Let $\varepsilon = (1 + \sqrt{p/n_1})^ 4 - 1$, which decreases as $n_1$ increases.
%	Let $\hat{A},\hat{B}$ be the global minimizer of $f(A, B)$.
%	With high probability over the randomness of the input,
%	the prediction loss of $\hat{\beta}_2^{\MTL} = \hat{B} \hat{A}_2$ for task two satisfies that
%	\begin{align}
%	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le
%	   \left|L(\hat{\beta}_2^{\MTL}) - \frac{2d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} -\frac{\sigma^2 p}{n_1 + n_2 - p}  \right|
%	\le \varepsilon \cdot \frac{2d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} +  \OO(p^{-c/2}).\label{cor_MTL_error}
%	%\left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] d^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \\
%	%& +C \left[(p^{-c_\varphi}+p^{-c_\infty/2})(\sigma^2 +d^2)+p^{-c_\infty}\kappa^2 + %\frac{d^4+\sigma^2 d^2}{\kappa^2}\right],\nonumber
%	 \end{align}
%	 with high probability for any fixed $c\in(0, \min(\frac{1}{4}, \delta,\frac{\varphi-4}{2\varphi}))$.
%	 {\color{red}[FY: the error also contains $p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) $, both of which cannot be omitted, because (i) there is no assumption on the upper bound of $\kappa^2$, and (ii) we do not necessarily have $c_\varphi<1/4$. We can decide how to present the result concisely (for instance we can impose an upper bound on $\kappa^2$ and that $c_\varphi<1/4$), but it needs to be correct.]}
	 \end{theorem}

\iffalse
 We need to state a result for Gaussian matrix ...... Consider
$$f (\al,n_1,n_2)= \frac1p\tr\left[\frac{1}{ (X_1^\top X_1 + \al \cdot X_2^\top X_2)^2} (X_1^\top X_1)^2\right].$$
In our case, we have $\al=1$, but we can handle more general $\al$. We introduce two parameters:
$$a= \al \frac{n_2}{n_1} \left( \frac{p}{n_1} + \frac{p}{n_2}- \frac{p}{n_1 }\cdot \frac{p}{n_2}\right),\quad b= \al \frac{n_2}{n_1}\left( 1- \frac{p}{n_2}\right) + \left( 1- \frac{p}{n_1}\right). $$
Then we define the following parameters:
\begin{align*}
x= \frac{-b+ \sqrt{b^2 + 4a}}{2a},\quad y= \left[ x^{-2} - \frac{p}{n_1}\left( 1+\frac{p}{n_1}x\right)^{-2}\right]^{-1},\quad \omega= \al\frac{n_2}{n_1} \left( 1 + \al \frac{p}{n_1}x\right)^{-1}.
\end{align*}
We have that
$$f(\al,n_1,n_2)= \frac{1 - 2\omega x + \omega^2 y}{ 1 - \frac{p}{n_2} \cdot \omega^2 y } +\oo(1)\quad \text{w.h.p.} $$
In the setting $\al=1$, both $\omega x$ and $\omega^2 y$ can be written in terms of only one parameter
%$$f(\al,n_1,n_2)= \frac{\left(u^2 -\frac{p}{n_1}\right) \left(1- 2\frac{n_2}{n_1}  u^{-1}\right) + \frac{n_2^2}{n_1^2} }{ u^2 -\frac{p}{n_1}\left(1 + \frac{n_2}{n_1}\right) } +\oo(1)\quad \text{w.h.p.} $$
%where
$$u: = x^{-1}\left(1+\frac{p}{n_1}x\right)=   \frac{ b+\sqrt{b^2+4a}}{2} +\frac{p}{n_1}   .$$
\fi

We believe the convergence rates $p^c/n_1$ and $p^{-1/2+c}$ in \eqref{Lvar_samplesize} and \eqref{Lbias_samplesize} are both sharp up to the $p^c$ factor. The proof of Theorem \ref{cor_MTL_loss} will be given in Appendix \ref{app_iso_cov} in the supplement \cite{MTL_suppl}. For the variance estimate in \eqref{Lvar_samplesize}, it is not necessary to assume the Gaussian distributions of the $Z^{(1)}$ and $Z^{(2)}$ entries. In fact, \eqref{Lvar_samplesize} is a special case of Theorem \ref{thm_main_RMT} below for the more general case with possibly different $\Sigma^{(1)}$ and $\Sigma^{(2)}$. On the other hand, the Gaussian assumption is needed in our current proof of the bias limit \eqref{Lbias_samplesize}. In the setting of Theorem \ref{cor_MTL_loss}, we can write
\begin{align*}
L_{\bias} (a)&=\bv_a^\top (Z^{(1)})^\top Z^{(1)} \frac1{ \left[a^2(Z^{(1)})^\top Z^{(1)}+ (Z^{(2)})^\top Z^{(2)}  \right]^{2}}(Z^{(1)})^\top Z^{(1)} \bv_a,
\end{align*}
where we abbreviate $\bv(a):=(\Sigma^{(1)})^{1/2}\left(a\beta^{(1)}- a^2\beta^{(2)}\right)$. By the rotational invariance of $(Z^{(1)})^\top Z^{(1)}$ and $(Z^{(2)})^\top Z^{(2)}$, we have that
\begin{align}\label{Lbias_idea}
L_{\bias} (a)&\approx \|\bv_a\|^2 \frac1p\bigtr{ [(Z^{(1)})^\top Z^{(1)}]^2 \frac1{ \left[a^2(Z^{(1)})^\top Z^{(1)}+ (Z^{(2)})^\top Z^{(2)}\right]^{2}}  },
\end{align}
up to a small error. Notice that we can write \eqref{Lbias_idea} into a simpler form
\begin{align}\nonumber %\label{Lbias_idea2}
L_{\bias} (a)&\approx \|\bv_a\|^2 \left. \frac{\dd }{\dd x}\right|_{x=0}\frac1p\bigtr{  \frac1{ a^2(Z^{(1)})^\top Z^{(1)}+ x[(Z^{(1)})^\top Z^{(1)}]^2+(Z^{(2)})^\top Z^{(2)} }  }.
\end{align}
It is well-known that the empirical spectral distributions (ESD) of $(Z^{(i)})^\top Z^{(i)}$, $i=1,2$, satisfy the famous Marchenko-Pastur (MP) law asymptotically \cite{MP}. From the MP law of $(Z^{(1)})^\top Z^{(1)}$, we can also derive the asymptotic ESD of $a^2(Z^{(1)})^\top Z^{(1)}+ x[(Z^{(1)})^\top Z^{(1)}]^2$ for any fixed $a\in \R$ and $x>0$. Due to the rotational invariance of multivariate Gaussian distributions,
%$a^2(Z^{(1)})^\top Z^{(1)}+ x[(Z^{(1)})^\top Z^{(1)}]^2$ and $(Z^{(2)})^\top Z^{(2)}$ are asymptotically freely independent from each other \cite{nica2006lectures}. Hence
the asymptotic ESD of $a^2(Z^{(1)})^\top Z^{(1)}+ x[(Z^{(1)})^\top Z^{(1)}]^2+(Z^{(2)})^\top Z^{(2)}$ is given by the free additive convolution (or free addition) of the asymptotic ESD of $a^2(Z^{(1)})^\top Z^{(1)}+ x[(Z^{(1)})^\top Z^{(1)}]^2$ and the MP law of $(Z^{(2)})^\top Z^{(2)}$\cite{nica2006lectures}. In particular, a sharp convergence estimate has been proved in \cite{BES_free1,BES_free2} for the free addition of two probability measures. We use that result to obtain a convergence estimate on
$$\frac1p\bigtr{  \frac1{ a^2(Z^{(1)})^\top Z^{(1)}+ x[(Z^{(1)})^\top Z^{(1)}]^2+(Z^{(2)})^\top Z^{(2)} }  }.$$
Taking derivative with respect to $x$ at $x=0$ gives the exact asymptotic limit of $L_{\bias}(a)$. We refer the reader to Appendix \ref{app_iso_cov} of the supplement \cite{MTL_suppl} for more details.

We believe that the above argument can be extended to the case without the Gaussian assumption. For example, instead of using the results in \cite{BES_free1,BES_free2} on free addition, we can use the sharp local laws on polynomials of random matrices in \cite{EKN_poly}. However, to apply the result in \cite{EKN_poly}, we need to check some difficult technical regularity conditions for our setting, which is not the focus of this work. Hence we do not pursue this direction in this paper.

\subsection{Proof overview}
For the rest of this section, we present an overview of the proof of Theorem \ref{thm_main_RMT}.
The central quantity of interest is the inverse of the sum of two sample covariance matrices.
We note that the variance equation $\tr[\Sigma^{(2)} \hat{\Sigma}^{-1}]$ is equal to $(n_1 + n_2)^{-1} \bigtr{W^{-1}}$, where $W$ is
\begin{align}\label{eigen2extra}
\frac{\Lambda U^\top (Z^{(1)})^\top Z^{(1)} U\Lambda  + V^\top (Z^{(2)})^\top Z^{(2)}V}{n_1 + n_2}.
\end{align}
Here $U\Lambda V^\top$ is defined as the SVD of $M$.
This formulation is helpful because we know that $(Z^{(1)})^{\top} Z^{(1)}$ and $(Z^{(2)})^{\top} Z^{(2)}$ are both sample covariance matrices with isotropic population covariance, and $U, V$ are both orthonormal matrices.
For example, if $Z^{(1)},Z^{(2)}$ are both Gaussian random matrices, by rotational invariance, $Z^{(1)} U, Z^{(2)}V$ are still Gaussian random matrices.

Our proof uses the Stieltjes transform or the resolvent method in random matrix theory.
We briefly describe the key ideas and refer the interested readers to classical texts such as  \cite{bai2009spectral,tao2012topics,erdos2017dynamical}.
For any probability measure $\mu$ supported on $[0,\infty)$, the Stieltjes transform of $\mu$ is a complex function defined as
$$m_\mu(z):= \int_0^\infty \frac{\dd\mu(x)}{x-z}, \text{ for any complex } z\in \C\setminus \set{0}.$$
Thus, the Stieltjes transform method reduces the study of a probability measure $\mu$ to the study of a complex function $m_\mu(z)$.

Let $\mu=p^{-1}\sum_{i} \delta_{\sigma_i}$ denote the empirical spectral distribution of $W$, where the $\sigma_i$'s are the eigenvalues of $W$ and $\delta_{\sigma_i}$ is the point mass measure at $\sigma_i$. Then it is easy to see that the Stieltjes transform of $\mu$ is equal to
\[ m_{\mu}(z) \define \frac{1}{p}\sum_{i=1}^p \frac{1}{\sigma_i - z}= p^{-1}\tr\left[(W-z\id)^{-1}\right]. \]
The above matrix $(W - z\id)^{-1}$ is known as $W$'s resolvent or Green's function.
We prove the convergence of $W$'s resolvent using the so-called ``local law'' with a sharp convergence rate \cite{isotropic,erdos2017dynamical,Anisotropic}.

We say that $(W-z\id)^{-1}$ converges to a deterministic $p\times p$ matrix limit $R(z)$ if for any sequence of deterministic unit vectors $v\in \R^p$,
$$v^\top \left[(W-z\id)^{-1}-R(z)\right]v\to 0\ \ \ \text{when $p$ goes to infinity.
}$$
To study $W$'s resolvent, we observe that $W$ is equal to $\AF\AF^{\top}$ for a $p$ by $n_1 + n_2$ matrix
\be\label{defn AF} \AF := (n_1+ n_2)^{-1/2} [\Lambda U^\top (Z^{(1)})^\top,V^\top (Z^{(2)})^\top]. \ee
Consider the following symmetric block matrix whose dimension is $p + n_1 + n_2$
\begin{equation}\label{linearize_block}
H \define \left( {\begin{array}{*{20}c}
0 & \AF  \\
\AF^{\top} & 0
\end{array}} \right).
\end{equation}
For this block matrix, we define its resolvent as
$$G(z) \define \left[H - \begin{pmatrix}z\id_{p\times p}&0\\ 0 & \id_{(n_1+n_2)\times (n_1+n_2)} \end{pmatrix}\right]^{-1},$$
for any complex value $z\in \mathbb C$.
Using Schur complement formula for the inverse of a block matrix, it is not hard to verify that
\begin{equation} \label{green2}
G(z) =  \left( {\begin{array}{*{20}c}
(W- z\id)^{-1} & (W - z\id)^{-1} \AF  \\
\AF^\top (W - z\id)^{-1} & z(\AF^\top \AF - z\id)^{-1}
\end{array}} \right).%
\end{equation}

\textbf{Variance asymptotic limit.}
In Theorem \ref{main_cor}, we will show that for $z$ in a small neighborhood around $0$, when $p$ goes to infinity, $G(z)$ converges to the following limit
\be \label{defn_piw}
\Gi(z) \define \begin{pmatrix} (a_{1}(z)\Lambda^2  +  (a_{2}(z)- z)\id_{p\times p})^{-1} & 0 & 0 \\ 0 & - \frac{n_1+n_2}{n_1} a_{1}(z)\id_{n_1\times n_1} & 0 \\ 0 & 0 & -\frac{n_1+n_2}{n_2}a_{2}(z)\id_{n_2\times n_2}  \end{pmatrix},\ee
where $a_1(z)$ and $a_2(z)$ are the unique solutions to the following self-consistent equations
\be\label{selfomega_a}
\begin{split}
&a_1(z) + a_2(z) = 1 - \frac{1}{n_1 + n_2} \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1(z) + a_2(z)}{\lambda_i^2 a_1(z) + a_2(z) - z}}, \\ %
&a_1(z) + \frac{1}{n_1 + n_2}\bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1(z)}{\lambda_i^2 a_1(z) + a_2(z) - z}} = \frac{n_1}{n_1 + n_2}.
\end{split}
\ee
The existence and uniqueness of solutions to the above system are shown in Lemma \ref{lem_mbehaviorw}.
Given this result, we now show that when $z = 0$, the matrix limit $\Gi(0)$ implies the variance limit shown in equation \eqref{lem_cov_shift_eq}.
First, we have that $a_1 = a_1(0)$ and $a_2 = a_2(0)$ since the equations in \eqref{selfomega_a} reduce to equation \eqref{eq_a12extra} when $z=0$.
Second, since $W^{-1}$ is the upper-left block matrix of $G(0)$, we have that $W^{-1}$ converges to $ (a_1\Lambda^2 + a_2\id)^{-1} $.
Using the fact that $\tr[\Sigma^{(2)} \hat{\Sigma}^{-1}] = (n_1 + n_2)^{-1}\bigtr{W^{-1}} $, we get that when $p$ goes to infinity, %
\begin{align*}
\bigtr{\Sigma^{(2)} \hat{\Sigma}} \rightarrow \frac{1}{n_1+n_2}\bigtr{(a_1 \Lambda^2 + a_2\id)^{-1}} &= \frac1{n_1+n_2}\bigtr{(a_1 M^{\top}M + a_2 \id)^{-1}} \\
&=\frac{1}{n_1+n_2} \bigtr{\Sigma^{(2)} (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}},
\end{align*}
where we note that $M^\top M = (\Sigma^{(2)})^{-1/2} \Sigma^{(1)} (\Sigma^{(2)})^{-1/2}$ and its SVD is equal to $V^{\top}\Lambda^2 V$.

\textbf{Bias asymptotic limit.}
For the bias limit in equation \eqref{lem_cov_derv_eq}, we show that it is governed by the derivative of $(W - z\id)^2$ with respect to $z$ at $z = 0$.
First, we can express the empirical bias term in equation \eqref{lem_cov_derv_eq} as %
\begin{align}\label{calculate G'}
(n_1 + n_2)^2 \hat{\Sigma}^{-1}\Sigma^{(2)}\hat{\Sigma}^{-1} = {\Sigma^{(2)}}^{-1/2} V W^{-2} V^{\top} {\Sigma^{(2)}}^{-1/2}.
\end{align}
Let $\cal G(z):=(W-z\id )^{-1}$ denote the resolvent of $W$.
Our key observation is that $\frac{\dd{\cal G(z)}}{\dd z} =  \cal G^2(z)$.
Hence, provided that the limit of $(W - z\id)^{-1}$ is $(a_1(z) \Lambda^2 + (a_2(z) - z) \id)^{-1}$ near $z = 0$, the limit of $\frac{\dd{\cal G(0)}}{\dd z}$ satisfies
\begin{align}\label{cal G'0}
\frac{\dd \cal G(0)}{\dd z} \to \frac{-\frac{\dd a_1(0)}{\dd z}\Lambda^2 - (\frac{\dd a_2(0)}{\dd z} - 1)\id}{(a_{1}(0)\Lambda^2 + a_{2}(0)\id_p)^2}.
\end{align}
To find the derivatives of $a_1(z)$ and $a_2(z)$, we take the derivatives on both sides of the system of equations \eqref{selfomega_a}.
Let $a_3 = - \frac{\dd a_1(0)}{\dd z}$ and $a_4 = - \frac{\dd a_2(0)}{\dd z}$.
One can verify that $a_3$ and $a_4$ satisfy the self-consistent equations in \eqref{eq_a34extra} (details omitted).
Applying equation \eqref{cal G'0} to equation \eqref{calculate G'}, we obtain the bias limit.

As a remark, in order for $\frac{\dd \cal G(z)}{\dd z}$ to stay close to its limit at $z = 0$, we not only need to find the limit of $\cal G(0)$, but also the limit of $\cal G(z)$ within a small neighborhood of $0$.
This is why we consider $W$'s resolvent for a general $z$ (as opposed to the Stieljes transform of its empirical spectral distribution discussed earlier).

\textbf{Schur complement and self-consistent equations.}
First, we consider the special case where $Z^{(1)}$ and $Z^{(2)}$ are both multivariate Gaussian random matrices.
By rotational invariance, we have that $Z^{(1)} U$ and $Z^{(2)} V$ are still multivariate Gaussian random matrices.
Next, we use the Schur complement formula to deal with the resolvent $G(z)$.
We show that $G(z)$'s diagonal entries satisfy a set of self-consistent equations in the limit, leading to equations in \eqref{selfomega_a}.
On the other hand, $G(z)$'s off-diagonal entries are approximately zero using standard concentration bounds.
Finally, we extend our result to general random matrices under the finite $\varphi$-th moment condition.
We prove an anisotropic local law using recent developments in random matrix theory \cite{erdos2017dynamical,Anisotropic}.
The proof of Theorem \ref{thm_main_RMT} is shown in Appendix \ref{appendix RMT}.

\input{examples}

\input{exp}

\input{multitask}

\section{Conclusions}\label{sec_conclude}

This work studied generalization properties of a widely used hard parameter sharing approach for multi-task learning.
We provided sharp bias-variance tradeoffs of HPS in high-dimensional linear regression.
Using these results, we analyzed how varying sample sizes and covariate shifts impact HPS, and rigorously explained several empirical phenomena such as negative transfer and covariate shift related to these dataset properties.
We validated our theory and conducted further studies on text classification tasks.
We describe several open questions for future work.
%First, it would be interesting to tighten our estimate in Corollary \ref{cor_MTL_loss}, which would extend the observation in Figure \ref{fig_size} to small $n_1$.
%Second, it would be interesting to extend our result to classification problems such as logistic regression.
First, our result in Corollary \ref{cor_MTL_loss} involves an error term that scales down with $n_1$.
Tightening this error bound requires showing the limit of $\normFro{({Z^{(1)}}^{\top} Z^{(1)} + {Z^{(2)}}^{\top} Z^{(2)})^{-1} {Z^{(1)}}^{\top} Z^{(1)}}^2$ for two isotropic sample covariance matrices.
This requires studying the asymptotic singular values distribution of the non-symmetric matrix $({Z^{(1)}}^{\top} Z^{(1)})^{-1}{Z^{(2)}}^{\top} Z^{(2)}+\id$, which is still an open problem in random matrix theory.
%FY: $+\id$ is very important and makes the problem very hard; otherwise the problem can be solved with current RMT methods.
The eigenvalue distribution of this matrix, which has been obtained in \cite{Fmatrix}, might be helpful towards resolving this problem.
%but its singular values will follow a different distribution since the matrix is not symmetric.
 %might require new techniques beyond the current ones in random matrix theory .%\HZ{to add}.
Second, it would be interesting to extend our results to classification problems.
Several recent work have made remarkable progress for logistic regression in the high-dimensional setting, e.g. \cite{sur2019modern}.
An interesting question is to study logistic regression in a multiple-sample setting.

\begin{supplement}
\textbf{Supplement to ``High-dimensional Asymptotics of Transferring from Different Domains in Linear Regression"}.
In the Internet Appendix \cite{MTL_suppl}, we provide the proofs of the technical results in Sections \ref{sec_HPS}-\ref{sec_diff}, including Lemma \ref{lem_HPS_loss}, Theorem \ref{thm_many_tasks}, Theorem \ref{cor_MTL_loss}, Proposition \ref{lem_hat_v}, Theorem \ref{thm_main_RMT} and Proposition \ref{prop_main_RMT}.
%and give the technical proofs of the main theorems, Theorems \ref{thm_regularbm}, \ref{thm_twgram}, \ref{thm_twgraph} and  \ref{thm_twgraph_sparse}.
\end{supplement}

\bibliographystyle{imsart-number}
\bibliography{rf,ref_mtl}

\end{document}