### 2021-06-12 revision_journal/main_arxiv.tex

\documentclass{article}

\def\shownotes{1}

\usepackage{natbib}
\setcitestyle{authoryear,open={(},close={)}}
\renewcommand{\cite}{\citep}
\newcommand{\yell}[1]{{\color{red}#1}}
\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}
\usepackage{authblk}
\usepackage[margin=1in]{geometry}
%\usepackage{hyperref}

\usepackage{macro_math}
\usepackage[us,12hr]{datetime} % `us' makes \today behave as usual in TeX/LaTeX
\usepackage{graphicx, subcaption}
\usepackage[export]{adjustbox}
\usepackage{subfloat}
\usepackage[T1]{fontenc}    %
\usepackage{nicefrac}       %
\usepackage{microtype}      %
\usepackage[hidelinks]{hyperref}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}}
\usepackage{macro_fan}
\usepackage{stmaryrd}
\usepackage[utf8]{inputenc} %
\usepackage{caption,subcaption}

%\newcommand{\citet}{\cite}

\begin{document}

	\iffalse
	\vspace{0.15in}
	\author[1]{Hongyang R. Zhang\thanks{HZ and FY contributed equally to the theoretical results. SW conducted the experimental analysis. Correspondence to hrzhang@northeastern.edu or fyang75@wharton.upenn.edu.}}
	\author[2]{Fan Yang\samethanks}
	\author[3]{Sen Wu\samethanks}
	\author[2]{Weijie J. Su}
	\author[3]{Christopher R\'e}
	\vspace{0.15in}
	\affil[1]{Khoury College of Computer Sciences, Northeastern University}
	\affil[2]{Department of Statistics, The Wharton School, University of Pennsylvania}
	\affil[3]{Department of Computer Science, Stanford University}
	\fi
	\title{Precise Asymptotics of Transferring from Different Regression in High Dimensions}
	\maketitle
	\input{abstract}
	\input{intro}
	\input{results}
	\input{examples}
	\input{multitask}
	\input{exp}
	\input{conclusions}
	\bibliographystyle{plainnat}
	\bibliography{rf,ref_mtl}

\end{document}