### 2021-07-22 revision_journal/main_arxiv.tex

\documentclass{article}

\def\shownotes{1}
\setlength{\columnsep}{0.35in}
\usepackage{natbib}
\setcitestyle{authoryear,open={(},close={)}}
\renewcommand{\cite}{\citep}
\newcommand{\yell}[1]{{\color{red}#1}}
\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}
%\usepackage{authblk}
\usepackage[margin=1in]{geometry}
%\usepackage{hyperref}

\usepackage{macro_math}
\usepackage[us,12hr]{datetime} % `us' makes \today behave as usual in TeX/LaTeX
\usepackage{graphicx, subcaption}
\usepackage[export]{adjustbox}
\usepackage{subfloat}
\usepackage[T1]{fontenc}    %
\usepackage{nicefrac}       %
\usepackage{microtype}      %
\usepackage[hidelinks]{hyperref}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}}
\usepackage{macro_fan}
\usepackage{stmaryrd}
\usepackage[utf8]{inputenc} %
\usepackage{caption,subcaption}

%\newcommand{\citet}{\cite}

\begin{document}

	\title{Estimating Information Transfer from Different Sources in High-dimensional Linear Regression}

    \author{Fan Yang\thanks{University of Pennsylvania. fyang75@wharton.upenn.edu}\and Hongyang R. Zhang \thanks{Northeastern University. ho.zhang@northeastern.edu. Part of this work was done while the author was at UPenn.}\and Sen Wu\thanks{Stanford University. senwu@stanford.edu}\and Weijie J. Su \thanks{University of Pennsylvania. suw@wharton.upenn.edu}\and Christopher R\'e\thanks{Stanford University. chrismre@stanford.edu}}

	\maketitle
	\input{abstract}
	\input{intro}
	\input{setup}
	\input{results}
	\input{exp}
    \input{multitask.tex}
	\input{conclusions}
	\bibliographystyle{plainnat}
	\bibliography{rf,rf_tl,ref_mtl}

\end{document}